---
title: "DEG with DESeq2"
output: html_notebook
---
First set the root directory as the working directory.
By default getwd() is the directory of the markdown file.
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = dirname(getwd()))
```

```{r warning=FALSE, message=FALSE}
library(DESeq2)
library(tidyverse)
library(ggplot2)
library(viridis)
library(ggh4x)
library(patchwork)
library(ggrepel)
library(UpSetR)
library(clusterProfiler)
library(org.Mmu.eg.db)
library(openxlsx)
source("scripts/translate_mmul-human.R")
```

load the dds object
```{r}
file_name <- list.files("data/dds/")
# only rds
file_name <- file_name[grepl("rds", file_name)]
# take the most recent file
file_name <- file_name[length(file_name)]
dds <- readRDS(paste0("data/dds/", file_name))
```

Create design matrices
```{r}
sample_metadata <- colData(dds)
# full design
design <- model.matrix(~ Batch + Treatment:Animal.n + Treatment + Week_post_FHT + Treatment:Week_post_FHT, sample_metadata)
# remove duplicates
idx <- which(apply(design, 2, function(x) all(x==0)))
if (length(idx) > 0) design <- design[,-idx]

# reduced designs
design_red_TreTime <- model.matrix(~  Batch + Treatment + Week_post_FHT + Treatment:Animal.n, sample_metadata)

# remove duplicates
idx <- which(apply(design_red_TreTime, 2, function(x) all(x==0)))
if (length(idx) > 0) design_red_TreTime <- design_red_TreTime[,-idx]
```

Run DESeq2

```{r}
dds <- DESeq(dds,
             full = design # full design table
)
res <- results(dds)
res
# save dds object with date
# create folder if it does not exist
if (!dir.exists("results/dds")) dir.create("results/dds")
saveRDS(dds, paste0("results/dds/dds", format(Sys.Date(), "%Y%m%d"), ".rds"))
```


Explore results.

```{r}
list_res <- list()
for (res_name in resultsNames(dds)) {
  res <- results(dds, name = res_name)
  list_res[[res_name]] <- results(dds, name = res_name, tidy = TRUE)
  # res
  print(res_name)
  summary(res)
}
```
save tables to excel
```{r}
# if folder does not exist, create it
if (!dir.exists("results/tables/02_DEG")) dir.create("results/tables/02_DEG")
write.xlsx(list_res, "results/tables/02_DEG/results_DESeq2.xlsx")
```


plot p-values

```{r}
# collapse result tables from list and add column with the name of the result
all_res <- bind_rows(list_res, .id = "result_name")
# plot p-values
p_value <- all_res%>%
        drop_na(padj) %>%
        ggplot(aes(x = pvalue) )+
  geom_histogram(aes(y=after_stat(count/sum(count))), breaks = seq(0, 1, 0.05),
                 color = "black", fill = "grey") +
  # Remove space between x-axis and min(y)
  scale_y_continuous(expand = expansion(c(0, 0.05))) +
  facet_wrap(vars(result_name), scales="free_y") + # separate plots
  theme_bw(base_size = 12)
p_value
```

# save plot
```{r}
pdf("results/plot/02_DEG/p_value_histogram.pdf", width = 15, height = 10)
p_value
dev.off()
```




# upset plots with results
All results

```{r}
# get significant genes
result_names <- resultsNames(dds)
# result_names <- result_names[!grepl("Animal", result_names)]
# result_names <- result_names[!grepl("Batch", result_names)]
sig_genes <- list()
for (res_name in result_names) {
  res <- results(dds, name = res_name, tidy = TRUE)
  if (length(res) == 0) next
  sig_genes[[res_name]] <- na.omit(res[res$padj < 0.05,]$row)
}
# remove Intercept
sig_genes <- sig_genes[-1]

upset <-  upset(fromList(sig_genes), sets.bar.color = "#566be9", main.bar.color = "#566be9", order.by = "freq",
      nsets = length(result_names)-1)
print(upset)
pdf("results/plot/02_DEG/upset.pdf", width = 20, height = 12, onefile=F)
upset
dev.off()
```


 just E2 effects
```{r}
# get significant genes
result_names <- resultsNames(dds)
result_names <- result_names[grepl("E2.W", result_names)]

sig_genes <- list()
for (res_name in result_names) {
  res <- results(dds, name = res_name, tidy = TRUE)
   if (length(res) == 0) next
  sig_genes[[res_name]] <- na.omit(res[res$padj < 0.05,]$row)
}


upset_E2<- upset(fromList(sig_genes), sets.bar.color = "#566be9", main.bar.color = "#566be9", order.by = "freq",
      nsets = length(result_names)-1, text.scale = 2)
print(upset_E2)

pdf("results/plot/02_DEG/upset_E2_effects.pdf", onefile = F)
upset_E2
dev.off()
```

vaccination effects
```{r}
# get significant genes
result_names <- resultsNames(dds)
result_names <- result_names[grepl("^W", result_names)]

sig_genes <- list()
for (res_name in result_names) {
  res <- results(dds, name = res_name, tidy = TRUE)
   if (length(res) == 0) next
  sig_genes[[res_name]] <- na.omit(res[res$padj < 0.05,]$row)
}


upset_Vax <- upset(fromList(sig_genes), sets.bar.color = "#566be9", main.bar.color = "#566be9", order.by = "freq",
      nsets = length(result_names)-1, text.scale = 2)
print(upset_Vax)
pdf("results/plot/02_DEG/upset_vax_effects.pdf", onefile=F)
upset_Vax
dev.off()
```

# Make heatmaps

Get vsd count matrix for visualization
```{r}
# transform to vst
vsd <- vst(dds, blind=FALSE)
# make dataframae with counts
vsd <- assay(vsd)
vsd <- as.data.frame(vsd)
vsd_scaled <- t(scale(t(vsd)))
# save vsd

```

For every week, make a heatmap with all the significant genes for the interaction
```{r warning=FALSE, message=FALSE}
result_names <- resultsNames(dds)
# remove intercept
result_names <- result_names[-1]
result_names <- result_names[grepl("Week", result_names)]


heatmap_list <- list()
for (res_name in result_names) {
  heatmap_list[[res_name]] <- get_heatmap_res(res_name, vsd_scaled, dds, trans_table = trans_table)
}

pdf("results/plot/02_DEG/heatmap_weeks.pdf", width = 15, height = 10)
heatmap_list
dev.off()
```

all of them
```{r} warning=FALSE, message=FALSE}
result_names <- resultsNames(dds)
# remove intercept
result_names <- result_names[-1]


heatmap_list <- list()
for (res_name in result_names) {
  heatmap_list[[res_name]] <- get_heatmap_res(res_name, vsd_scaled, dds, trans_table = trans_table)
}

pdf("results/plot/02_DEG/heatmap_all.pdf", width = 15, height = 10)
heatmap_list
dev.off()
```



# Volcano plots

```{r}
result_names <- resultsNames(dds)
# remove intercept
result_names <- result_names[-1]

# get max log2 fold change across all of them for xlim, centered at 0
max_lfc <- lapply(result_names, function(x) max(abs(results(dds, name = x, tidy = TRUE)$log2FoldChange)))
max_lfc <- max(as.numeric(unlist(max_lfc)))

volcano_list <- lapply(result_names, function(x) get_vln_plot(dds, x, trans_table, xlim=c(-max_lfc, max_lfc)))
pdf("results/plot/02_DEG/volcano.pdf", width = 15, height = 10)
volcano_list
dev.off()
```

# Tables
save results table
```{r}
result_names <- resultsNames(dds)
# remove intercept
result_names <- result_names[-1]
# save each table as sheet of excel
library(openxlsx)
result_table_list <- list()
for (res_name in result_names) {
  res <- results(dds, name = res_name, tidy = TRUE)
    result_table_list[[res_name]] <- res
}
write.xlsx(result_table_list, "results/tables/02_DEG/results_DESeq2.xlsx")
```


