---
title: "Olink analysis for NHP FHT"
output: html_notebook
---

Based on Camila's code

```{r setup, include=FALSE}
# set root dir for knitr
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

```


# Loading
## load libraries
```{r}
library(readxl)
library(OlinkAnalyze)
library(stringr)
library(tidyverse)
library(reshape2)
library(MuMIn)
library(cowplot)
library(lme4)
library(ggthemes)
library(parameters) # this should be installed
library(emmeans)
library(patchwork)
```

## load data

```{r}
meta <- read_excel("data/olink/Olink sample information.xlsx")
meta$Sample <- sapply(str_split(meta$SampleID, "-"), "[", 1)
Olink <- read_NPX(filename = "data/olink/AB00000321_NPX.xlsx")

# some typo in Rh-16 w27, annotated as placebo but must be E2
meta$Treatment[meta$SampleID %in% "Rh-16-27"] <- "E2"
```

# Explore data

## Check for missing values
```{r}
meta %>%
  ggplot(aes(x=Week, y=Sample, group=Treatment, color=Treatment)) + geom_point() +
  scale_x_continuous(breaks = c(0,7,13,18,19,23,27)) +
  ggtitle("Olink sample information") +
        facet_grid(rows=vars(Treatment), scales="free_y")
```


### filtering down to only useful samples and information

```{r}
# cast df
olink <- Olink[,c(1,5,12)] #Sample ID, Assay (the feature), NPX
olink <- reshape2::dcast(olink, SampleID ~ factor(Assay, levels = unique(olink$Assay)), fun.aggregate=function(i) mean(i, na.rm=TRUE))
rownames(olink) <- olink$SampleID
olink <- olink  %>% filter(SampleID %in% meta$SampleID) #remove plate controls
olink <- olink[,-1]
```

### Filtering of PPs with variance
```{r}
### Filtering of PPs with variance ###
#sapply(olink, function(i) sd(i, na.rm=T)/mean(i, na.rm=T) *100) %>% sort()
#remove for low variation: IL2, IL33, Beta-NGF, LIF, IL-2RB, IL-20, FGF-5, IL-24
PPvar <- sapply(olink, var) %>% sort()
PPvar
olink_var <- olink %>% dplyr::select(names(PPvar[PPvar > 0.1])) #77 PPs

m_olink <- merge(olink_var, meta, by.x="row.names", by.y="SampleID")

m_olink$Treatment <- factor(m_olink$Treatment, levels = c("Placebo", "E2"))
m_olink$Week <- factor(m_olink$Week)

colnames(m_olink) <- str_remove(colnames(m_olink), "-")
colnames(m_olink) <- str_remove(colnames(m_olink), " ")
colnames(m_olink) <- gsub("-","", colnames(m_olink))
colnames(m_olink)[86] <- "X4EBP1"
```

# define main function
```{r}
run_mem <- function(varlist, dataframe, cov1 = NULL, cov2 = NULL, cov3 = NULL, cov4 = NULL, random_eff) {

  # make list of formulas for each PP
  MEM_PP <- lapply(varlist, function(x) {
    # add the specified covariates
    formula_terms <- c(cov1, cov2, cov3, cov4)
    formula_terms <- formula_terms[!is.null(formula_terms)]
    # make formula string
    formula_str <- paste(x, "~", paste(formula_terms, collapse = "+"), "+", "(1|", random_eff, ")")
    # run model, if it fails return NULL
    mod2 <- try(lmer(formula_str, data = dataframe, na.action = na.exclude))
    if (inherits(mod2, "try-error")) {
      return(NULL) # return NULL if the model can't be fitted
    } else {
      return(mod2) # return the model if it can be fitted
    }
  })

  # Model check
  MEM_PP_1 <- as.list(MEM_PP)
  MEM_PP_1[sapply(MEM_PP_1, is.null)] <- NULL # remove NULLs, i.e. models that couldn't be fitted

  # analyte
  Analyte <- lapply(MEM_PP_1, function(f) summary(f)$call[2]) # get the analyte name
  Analyte <- sapply(str_split(string = Analyte, pattern = "~"), `[`, 1)
  Analyte <- gsub(pattern = "\\(", replacement = "", Analyte)
  Analyte <- gsub(pattern = "\\ ", replacement = "", Analyte)
  Analyte <- Analyte %>% as.data.frame()
  colnames(Analyte) <- "Analyte"

  # coef
  coefs <- lapply(MEM_PP_1, function(f) summary(f)$coefficients[,1])
  coefs <- as.data.frame(do.call(rbind, coefs))
  colnames(coefs) <- paste("estimate", colnames(coefs),sep="")

  #pvalue
  pvals <- lapply(MEM_PP_1, function(f) parameters::p_value(f, method = "wald",)[,2])
  pvals_name <- lapply(MEM_PP_1, function(f) parameters::p_value(f, method = "wald",)[,1])[1] %>% unlist()
  pvals <- as.data.frame(do.call(rbind, pvals))
  colnames(pvals) <- paste("pValue_", pvals_name,sep="")

  # Median residuals
  Amed <- lapply(MEM_PP_1, function(f) summary(f)$residuals) #get residuals
  Amed_na <- lapply(Amed, function(f) na.exclude(f)) #exclude NAs from residuals
  Amed_calc <- lapply(Amed_na, function(f) median(f)) #calculate median
  Amed_calc <- as.data.frame(do.call(rbind, Amed_calc))
  colnames(Amed_calc) <- "residuals"

  # R^2
  r2 <- lapply(MEM_PP_1, function(f) r.squaredGLMM(f)[2])
  r2 <- as.data.frame(do.call(rbind, r2))
  colnames(r2) <- "Rsq"

  Adf <- cbind(Analyte, Amed_calc, r2, coefs, pvals)
  return(Adf)
}
```


# run MEM
model 4, week:treatment interaction (week as factor)
```{r}

MEM_PP_4 <- run_mem(varlist = colnames(m_olink)[2:86],
                    dataframe = m_olink,
                    cov1 = "Week",
                    cov2 = "Treatment",
                    cov3 = "Week*Treatment",
                    random_eff = "Sample") # animal as random effect
rownames(MEM_PP_4) <- MEM_PP_4$Analyte
```

some boundary (singular) fit: see help('isSingular')
this means that for some analytes the model is not able to fit the data, and the model is singular. This is likely due to the fact that the data is not enough to fit the model. This is a common problem in mixed models, and it is not a problem per se. However, it is important to be aware of this, and to interpret the results with caution.



## add FDR correction to df: Create new columns with _fdr extension

```{r}
colnames(MEM_PP_4)
selected_columns <- MEM_PP_4[, 18:31] # seelct the ones that are pvalues
pvals_corrected <- lapply(selected_columns, function(col) p.adjust(col, method = "BH"))
for (i in 1:ncol(selected_columns)) {
  col_name <- colnames(selected_columns)[i]
  new_col_name <- paste0(col_name, "_fdr")
  MEM_PP_4[[new_col_name]] <- pvals_corrected[[i]]
}
```


## significant analytes
```{r}
# Get analyte names for each column with FDR < 0.1
significant_analytes_list <- lapply(pvals_corrected, function(fdr_values) {
  analytes <- rownames(selected_columns)[fdr_values <= 0.1]
  return(analytes)
})

significant_analytes_list
significant_analytes_list %>% names()

analytes_time <- significant_analytes_list[2:7] %>% unlist() %>% as.character() %>% unique()
analytes_E2 <- significant_analytes_list$pValue_TreatmentE2
analytes_timeE2interaction <- significant_analytes_list[9:14] %>% unlist() %>% as.character() %>% unique()

```




# visualizatoin
MCP4
```{r}
m_olink %>%
  ggplot(aes(x=Week, y=MCP4)) +
        geom_boxplot(aes(fill=Treatment), alpha=0.25, position=position_dodge(width=0.75)) +
                 geom_point(aes(color=Treatment), position = position_dodge(width = 0.75)) +
  ggtitle("MCP4")
  #facet_grid(rows=vars(Treatment), scales="free_y")
```


## heatmap for significant time effects:
```{r}
m_olink %>% colnames()
scaled_m_olink <- m_olink
scaled_m_olink[,2:86] <- sapply(scaled_m_olink[,2:86], scale)

# plot all PP changing with only vaccination timepoints
estimates_order19 <- MEM_PP_4 %>% filter(Analyte %in% unique(c(significant_analytes_list$pValue_Week19, significant_analytes_list$pValue_Week23, significant_analytes_list$pValue_Week27))) %>% arrange(estimateWeek19) %>% pull(Analyte)

# make plot
Fig1_vaccine_effects_olink <- scaled_m_olink %>%
  select(all_of(c("Row.names", "Treatment", "Week", "Sample",
                  unique(c(significant_analytes_list$pValue_Week19, significant_analytes_list$pValue_Week23, significant_analytes_list$pValue_Week27)) ))) %>% #colnames()
  pivot_longer(cols = 5:54) %>%
  mutate(name = factor(name, levels = estimates_order19),
         value = as.numeric(value)) %>%
  ggplot(aes(x = Sample, y = name, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  theme_minimal() +
  labs(x = "Week", y = "Analytes", fill = NULL) +
  geom_vline(xintercept = 18, linetype = "dashed") +
  geom_vline(xintercept = 22, linetype = "dashed") +
  facet_wrap(~Week, ncol = 7) + theme(axis.text.x = element_blank()) + labs(y=NULL)
```


## plot PP changing with week 13
```{r}
scaled_m_olink %>%
  #m_olink %>%
  select(all_of(c("Row.names", "Treatment", "Week", "Sample", significant_analytes_list$pValue_Week13))) %>% #colnames()
  pivot_longer(cols = 5:30) %>%
  mutate(value = as.numeric(value)) %>%
  ggplot(aes(x = Sample, y = name, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  theme_minimal() +
  labs(x = "Week", y = "Analytes", fill = NULL) +
  geom_vline(xintercept = 18, linetype = "dashed") +
  geom_vline(xintercept = 22, linetype = "dashed") +
  facet_wrap(~Week, ncol = 7) + theme(axis.text.x = element_blank()) + labs(y=NULL, title = "changing at week 13")

```



## changes associated to E2 treatment

maybe it is better to do the mean and se per week and treatment, instead of geom_smooth loess
```{r}
inter_w18 <- significant_analytes_list$`pValue_Week18:TreatmentE2`

scaled_m_olink %>%
  select(all_of(c("Row.names", "Treatment", "Week", "Sample", inter_w18))) %>% #colnames()
  pivot_longer(cols = 5:7) %>%
  ggplot(aes(x=as.numeric(as.character(Week)), y=value, group=as.numeric(as.character(Week)), color=Treatment)) +
  geom_line(aes(group=Sample)) + theme_classic() + geom_vline(xintercept = 18, linetype="dashed") + geom_vline(xintercept = 22, linetype="dashed") +

        scale_x_continuous(breaks = c(0,7,13,18,19,23,27)) +
  facet_wrap(~name, ncol=8, scales="free") + labs(y=NULL) +
  scale_color_manual(values = c("E2" = "#8A69D4", "Placebo" = "#FFD635")) +
  scale_fill_manual(values = c("E2" = "#8A69D4", "Placebo" = "#FFD635")) +
        geom_smooth(aes(group=Treatment, fill=Treatment), method = "loess", se = T, linetype = "dashed", size = 0.5)

inter_vacc <- c(significant_analytes_list$`pValue_Week19:TreatmentE2`, significant_analytes_list$`pValue_Week23:TreatmentE2`)

scaled_m_olink %>%
  select(all_of(c("Row.names", "Treatment", "Week", "Sample", unique(c(inter_w18,inter_vacc)) ))) %>% #colnames()
  pivot_longer(cols = 5:8) %>%
  mutate(name=factor(name, levels = c("FGF19", "CCL19", "MCP4", "CST5"))) %>%
  ggplot(aes(x=as.numeric(as.character(Week)), y=value, group=as.numeric(as.character(Week)), color=Treatment)) +
  geom_line(aes(group=Sample)) + theme_classic() + geom_vline(xintercept = 18, linetype="dashed") + geom_vline(xintercept = 22, linetype="dashed") +
  scale_color_manual(values = c("E2" = "#8A69D4", "Placebo" = "#FFD635")) +
  scale_fill_manual(values = c("E2" = "#8A69D4", "Placebo" = "#FFD635")) +
        scale_x_continuous(breaks = c(0,7,13,18,19,23,27)) +
  facet_wrap(~name, ncol=8, scales="free") + labs(y="scaled NPX", x="Weeks") +
        geom_smooth(aes(group=Treatment, fill=Treatment), method = "loess", se = T, linetype = "dashed", size = 0.5)

significant_line_plot <- m_olink %>%
  select(all_of(c("Row.names", "Treatment", "Week", "Sample", unique(c(inter_w18,inter_vacc)) ))) %>% #colnames()
  pivot_longer(cols = 5:8) %>%
  mutate(name=factor(name, levels = c("FGF19", "CCL19", "MCP4", "CST5"))) %>%
        # display the change
        group_by(Sample, name) %>%
        mutate(value = value- value[Week == 0]) %>%
        # calculate mean and sd per treatment
    group_by(Treatment, name, Week) %>%
        mutate(mean_value = mean(value, na.rm = T), sd_value = sd(value, na.rm = T)) %>% ungroup() %>%

         ggplot(aes(x=as.numeric(as.character(Week)), y=mean_value,  color=Treatment)) +
        geom_line(aes(group=Sample, y =value)) + theme_classic() +
        geom_vline(xintercept = c(18, 22), linetype="dashed") +
        scale_color_manual(values = c("E2" = "#8A69D4", "Placebo" = "#FFD635")) +
  scale_fill_manual(values = c("E2" = "#8A69D4", "Placebo" = "#FFD635")) +
        scale_x_continuous(breaks = c(0,7,13,18,19,23,27), labels = c("0", "7", "13", "", "19", "23", "27")) +
  facet_wrap(~name, ncol=8) +
        labs(y= expression(paste(Delta, "NPX")), x="Weeks") +
geom_ribbon(aes(ymin = mean_value - sd_value, ymax = mean_value + sd_value, fill=Treatment, color=NULL), alpha = 0.2) +
        theme_minimal() +
		theme(panel.grid.minor = element_blank())


significant_analytes_list
significant_line_plot
# save plot
ggsave("results/plot/olink/significant_line_plot.pdf", significant_line_plot, width = 13, height = 2.5)
```


## visualize MEM results
```{r}
sigprots <- unique(c(MEM_PP_4$Analyte[MEM_PP_4$`pValue_Week7:TreatmentE2` <= 0.05],
                     MEM_PP_4$Analyte[MEM_PP_4$`pValue_Week13:TreatmentE2` <= 0.05],
                     MEM_PP_4$Analyte[MEM_PP_4$`pValue_Week18:TreatmentE2` <= 0.05],
                     MEM_PP_4$Analyte[MEM_PP_4$`pValue_Week19:TreatmentE2` <= 0.05],
                     MEM_PP_4$Analyte[MEM_PP_4$`pValue_Week23:TreatmentE2` <= 0.05],
                     MEM_PP_4$Analyte[MEM_PP_4$`pValue_Week27:TreatmentE2` <= 0.05]
))
memplotorder <- MEM_PP_4 %>% filter(Analyte %in% sigprots) %>% arrange(`estimateWeek18:TreatmentE2`) #%>% pull(Analyte)
memplotorder$`pValue_Week7:TreatmentE2_fdr`

```
```{r}
A <- memplotorder %>%
  mutate(Analyte=factor(Analyte, levels = memplotorder$Analyte)) %>%
  ggplot(aes(x=`estimateWeek7:TreatmentE2`, y=Analyte)) +
  geom_point(aes(color= `pValue_Week7:TreatmentE2_fdr`<=0.1, size=abs(`estimateWeek7:TreatmentE2`))) + #
  geom_vline(xintercept = 0, alpha=0.3, linetype="dotted") + scale_color_manual(values = c("gray", "red")) + scale_x_continuous(limits = c(-3, 3)) +
  labs(x="Estimate", y=NULL, size=NULL, title = "Week 7")  + theme_classic() + theme(legend.position = "none")#+ guides(color=guide_legend(title="FDR<0.1")) #+

B <- memplotorder %>%
  mutate(Analyte=factor(Analyte, levels = memplotorder$Analyte)) %>%
  ggplot(aes(x=`estimateWeek13:TreatmentE2`, y=Analyte)) +
  geom_point(aes(color= `pValue_Week13:TreatmentE2_fdr`<=0.1, size=abs(`estimateWeek13:TreatmentE2`))) + #
  geom_vline(xintercept = 0, alpha=0.3, linetype="dotted") + scale_color_manual(values = c("gray", "red")) + scale_x_continuous(limits = c(-3, 3)) +
  labs(x="Estimate", y=NULL, size=NULL, title = "Week 13")  + theme_classic() + theme(legend.position = "none")#+ guides(color=guide_legend(title="FDR<0.1")) #+

C <- memplotorder %>%
  mutate(Analyte=factor(Analyte, levels = memplotorder$Analyte)) %>%
  ggplot(aes(x=`estimateWeek18:TreatmentE2`, y=Analyte)) +
  geom_point(aes(color= `pValue_Week18:TreatmentE2_fdr`<=0.1, size=abs(`estimateWeek18:TreatmentE2`))) + #
  geom_vline(xintercept = 0, alpha=0.3, linetype="dotted") + scale_color_manual(values = c("gray", "red")) + scale_x_continuous(limits = c(-3, 3)) +
  labs(x="Estimate", y=NULL, size=NULL, title = "Week 18")  + theme_classic() + theme(legend.position = "none")#+ guides(color=guide_legend(title="FDR<0.1")) #+

D <- memplotorder %>%
  mutate(Analyte=factor(Analyte, levels = memplotorder$Analyte)) %>%
  ggplot(aes(x=`estimateWeek19:TreatmentE2`, y=Analyte)) +
  geom_point(aes(color= `pValue_Week19:TreatmentE2_fdr`<=0.1, size=abs(`estimateWeek19:TreatmentE2`))) + #
  geom_vline(xintercept = 0, alpha=0.3, linetype="dotted") + scale_color_manual(values = c("gray", "red")) + scale_x_continuous(limits = c(-3, 3)) +
  labs(x="Estimate", y=NULL, size=NULL, title = "Week 19")  + theme_classic() + theme(legend.position = "none") #+ guides(color=guide_legend(title="FDR<0.1")) #+

E <- memplotorder %>%
  mutate(Analyte=factor(Analyte, levels = memplotorder$Analyte)) %>%
  ggplot(aes(x=`estimateWeek23:TreatmentE2`, y=Analyte)) +
  geom_point(aes(color= `pValue_Week23:TreatmentE2_fdr`<=0.1, size=abs(`estimateWeek23:TreatmentE2`))) + #
  geom_vline(xintercept = 0, alpha=0.3, linetype="dotted") + scale_color_manual(values = c("gray", "red")) + scale_x_continuous(limits = c(-3, 3)) +
  labs(x="Estimate", y=NULL, size=NULL, title = "Week 23", color="FDR<0.1")  + theme_classic() + theme(legend.position = "none") #+ guides(color=guide_legend(title="FDR<0.1")) #+

G <- memplotorder %>%
  mutate(Analyte=factor(Analyte, levels = memplotorder$Analyte)) %>%
  ggplot(aes(x=`estimateWeek27:TreatmentE2`, y=Analyte)) +
  geom_point(aes(color= `pValue_Week27:TreatmentE2_fdr`<=0.1, size=abs(`estimateWeek27:TreatmentE2`))) + #
  geom_vline(xintercept = 0, alpha=0.3, linetype="dotted") + scale_color_manual(values = c("gray", "red")) + scale_x_continuous(limits = c(-3, 3)) +
  labs(x="Estimate", y=NULL, size=NULL, title = "Week 27")  + theme_classic() + theme(legend.position = "none") #+ guides(color=guide_legend(title="FDR<0.1")) #+
legend <- get_legend(memplotorder %>%
                       mutate(Analyte=factor(Analyte, levels = memplotorder$Analyte)) %>%
                       ggplot(aes(x=`estimateWeek23:TreatmentE2`, y=Analyte)) +
                       geom_point(aes(color= `pValue_Week23:TreatmentE2_fdr`<=0.1, size=abs(`estimateWeek23:TreatmentE2`))) + #
                       scale_color_manual(values = c("gray", "red")) +
                       labs(x=NULL, y=NULL, size="Estimate", title = "Week 23", color="FDR<0.1")  + theme_classic() ) # + theme(legend.box.margin = margin(0, 0, 0, 12))  create some space to the left of the legend

MEM_res <- plot_grid(A,B,C,D,E,G,legend, ncol=7)
MEM_res
```


# models with CI

```{r}
# run model for every assay
assay_models <- m_olink %>%
# make long format
pivot_longer(cols = 2:86, names_to = "Assay", values_to="NPX") %>%
mutate(Assay = as.factor(Assay),
       NPX = as.numeric(NPX),
       Sample = as.factor(Sample),
       Treatment = as.factor(Treatment),
       Week = as.factor(Week)) %>%
        ungroup() %>%
group_by(Assay) %>%
summarise(model = list(lmer(NPX ~ Week*Treatment + (1|Sample),
          data=pick(everything()))))


```

confint

```{r}
# get the confidence intervals
assay_interaction <- assay_models %>%
        ungroup() %>%
rowwise(Assay) %>%
reframe(emmeans(model,  ~ Week*Treatment)  %>%
                contrast(interaction = c("pairwise", "pairwise"), adjust = "none", infer = TRUE) %>% as_tibble())  %>%
# filter the ones we need
filter(grepl("0 -", Week_pairwise))

```

adjust p_values
```{r}
assay_interaction <- assay_interaction %>%
        # adjusted per contrast
        group_by(Week_pairwise) %>%
mutate(p.adj = p.adjust(p.value, method = "fdr"))
```

## significant

fdr
```{r}
assay_interaction %>%
        filter(p.adj < 0.1)
```

## visualize

plot with effect sizes and confidence intervals
```{r}

# arrange by week 18
interaction_data_order <- assay_interaction %>%
        filter(Assay %in% assay_interaction$Assay[assay_interaction$p.value < 0.05]) %>%
        select(Assay, Week_pairwise, estimate) %>%
        pivot_wider(names_from = Week_pairwise, values_from = estimate) %>%
        arrange(`0 - 18`)

interaction_plot <- assay_interaction %>%
        filter(Assay %in% assay_interaction$Assay[assay_interaction$p.value < 0.05]) %>%
        mutate(Assay = factor(Assay, levels = interaction_data_order$Assay)) %>%
        ggplot(aes(x=estimate, y= Assay, color=p.adj <0.1)) +
        geom_errorbar(aes(xmin = lower.CL, xmax = upper.CL), width = 0.1) +
        geom_point(aes(size=abs(estimate))) +
        geom_vline(xintercept = 0, linetype = "twodash") +
        scale_color_manual(values = c("gray", "red")) +
        # limit of size
        scale_size_continuous(breaks = c(0.5, 1,  2), range = c(1, 5)) +
        theme_classic() +
        labs(x="Estimate", y=NULL, color="FDR<0.1",
             size = "Estimate size",
             title = "Estrogen treatments effects on serum inflammatory proteins",
        subtitle = "Mean estimate and 95% CI") +
        facet_wrap(~Week_pairwise,  nrow=1)

interaction_plot
```

heatmap of results

```{r}
 interaction_heatmap <- assay_interaction %>%
        filter(Assay %in% assay_interaction$Assay[assay_interaction$p.value < 0.05]) %>%
        mutate(Assay = factor(Assay, levels = interaction_data_order$Assay)) %>%
        ggplot(aes(x=Week_pairwise, y= Assay, fill = (lower.CL >0 | upper.CL <0 ))) +
        geom_tile() +
        geom_text(aes(label = paste(round(lower.CL,2), round(upper.CL, 2))), size = 4)
interaction_heatmap

```


# create main figure

```{r}
design <- "
A
A
A
A
A
A
A
B
B
B
"
main_figure <- interaction_plot   / significant_line_plot  +
        plot_layout(design = design) +
        plot_annotation(tag_levels = "A")  &
        theme_minimal() +
        theme(strip.background.x = element_rect(fill = "#f7f4ed", color="white"))
main_figure

# save
ggsave("results/plot/olink/olink_main_figure.pdf", main_figure, width = 10, height = 10)
```


# vertical version of lineplot

```{r}
vertical_significant_line_plot <- m_olink %>%
  select(all_of(c("Row.names", "Treatment", "Week", "Sample", unique(c(inter_w18,inter_vacc)) ))) %>% #colnames()
  pivot_longer(cols = 5:8) %>%
  mutate(name=factor(name, levels = c("FGF19", "CCL19", "MCP4", "CST5"))) %>%
        # display the change
        group_by(Sample, name) %>%
        mutate(value = value- value[Week == 0]) %>%
        # calculate mean and sd per treatment
    group_by(Treatment, name, Week) %>%
        mutate(mean_value = mean(value, na.rm = T), sd_value = sd(value, na.rm = T)) %>% ungroup() %>%

         ggplot(aes(x=as.numeric(as.character(Week)), y=mean_value,  color=Treatment)) +
        geom_line(aes(group=Sample, y =value)) + theme_classic() +
        geom_vline(xintercept = c(18, 22), linetype="dashed") +
        scale_color_manual(values = c("E2" = "#8A69D4", "Placebo" = "#FFD635")) +
  scale_fill_manual(values = c("E2" = "#8A69D4", "Placebo" = "#FFD635")) +
        scale_x_continuous(breaks = c(0,7,13,18,19,23,27), labels = c("0", "7", "13", "", "19", "23", "27")) +
  facet_wrap(~name, ncol=2) +
        labs(y= expression(paste(Delta, "NPX")), x="Weeks") +
geom_ribbon(aes(ymin = mean_value - sd_value, ymax = mean_value + sd_value, fill=Treatment, color=NULL), alpha = 0.2) +
        theme_minimal() +
		theme(panel.grid.minor = element_blank(),strip.background.x = element_rect(fill = "#f7f4ed", color="white"),
        strip.text.x.top = element_text(size = 22, face = "bold"))

# save
ggsave("results/plot/olink/significant_line_plot_vertical.pdf", vertical_significant_line_plot, width = 10, height = 10)
significant_line_plot
```